---
title: "Week 03 Assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This assignment is due by 6 pm on the evening of the 3rd class. It should be submitted via Github.

#### Exercises

We'll be exploring [fuel economy data](https://www.fueleconomy.gov/feg/download.shtml) from 1984-2021. You can review a data dictionary for the dataset [here](https://www.fueleconomy.gov/feg/ws/index.shtml#vehicle).

Utilize the following chunk to get started with your analysis.

```{r}
library(tidyverse)
library(here)

vehicles <- read_csv(here::here("Data", "vehicles.csv"),
                     col_tyfpes = cols(.default = "c"))
```

1. Create a reduced version of the dataset that includes the following fields:

-   `make`
-   `model`
-   `year`
-   `cylinders`
-   `drive`
-   `trany`
-   `evMotor`
-   `fuelCost08`
-   `highway08`
-   `city08`
-   `feScore`

Ensure that the `year`, fuelCost08`, `highway08`, `city08`, and `feScore` fields are coerced to a double.

Assign this to a new dataframe called `vehicles_sub`.

```{r}

```

2. Conduct an exploratory analysis of each field in the dataset. This should include:

- Demonstrating the structure of the dataset
- Counts of each character field (in addition to the `year` field)
- Distributions of each numeric field

Indicate anything interesting that you notice.

Hint: When finding counts of fields, sometimes it is helpful to sort the field from highest to lowest values.

```{r}

```


```{r}

```

```{r}

```

3. Some of the fields need to be cleaned to ease analysis. 

- Create a new field in `vehicles_sub` called `evMotor_flag` that indicates whether or not the vehicle has an electric motor.

```{r}

```


4. Create a graph that demonstrates how median city mileage has changed over time.

```{r}

```

5. Consider the top 5 car makes whose fuel economies have been evaluated most frequently. Create a graph that demonstrates how the city fuel mileage has changed over time for these makes.

```{r}

```
6. Create a question about the dataset that must be solved by using the `dplyr` verbs - `select`, `mutate`, `summarize`, `arrange`, etc.

-   Indicate what your question is. Consider a question that could be answered best by utilizing facets, multiple geometries, multiple aesthetics, or some combination of those.
-   Create a graph that assists in answering the question.
-   Write a short interpretation of the graph in order to answer your question.

