---
title: "Week 03 Class Examples"
author: "Ryan Harrington"
date: "10/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
movies <- read_csv("Data_Transformations/movies_clean.csv")
```

```{r}
glimpse(movies)
```

```{r}
select(movies, id, title, original_title)
```

```{r}
select(movies, -keywords)
```

```{r}
select(movies, id, original_title, genre_Action:genre_TV)
```

```{r}
movies_sel <-
  select(movies,
         id, contains("title"), original_language, budget, contains("vote"),
         starts_with("release"), where(is.double), starts_with("genre"))
```

```{r}
filter(movies_sel,
       genre_Action,
       vote_average >= 7.5)
```

```{r}
filter(movies_sel,
       original_language == "en" | original_language == "fr" | original_language == "es" | original_language == "it")
```

```{r}
filter(movies_sel,
       original_language %in% c("en", "fr", "es", "it"))
```

```{r}
romance_languages <- c("en", "fr", "es", "it")

filter(movies_sel,
       original_language %in% romance_languages)
```

```{r}
arrange(movies_sel, budget)
```

```{r}
arrange(movies_sel, -budget)
arrange(movies_sel, desc(budget))
```

```{r}
arrange(movies_sel, release_year, desc(budget))
```

```{r}
mutate(movies_sel,
       new_column_name = transformation)
```


```{r}
mutate(movies_sel,
       profit = revenue - budget,
       runtime_diff = runtime - mean(runtime, na.rm = T))
```

```{r}
summarize(movies_sel,
          budget_median = median(budget, na.rm = T),
          budget_mean = mean(budget, na.rm = TRUE))
```

```{r}
movies_group <- group_by(movies_sel, release_year)

summarize(movies_group,
          budget_median = median(budget, na.rm = T),
          budget_mean = mean(budget, na.rm = TRUE),
          movies_count = n())
```

What movie had the highest budget in 2001?

```{r}
arrange(filter(select(movies, id, original_title, budget, release_year), release_year == 2001), desc(budget))
```

```{r}
movies_selected <- select(movies_sel, id, original_title, budget, release_year)
movies_filtered <- filter(movies_selected, release_year == 2001)
movies_arranged <- arrange(movies_filtered, desc(budget))
```

```{r}
movies_sel %>% 
  select(id, original_title, budget, release_year) %>% 
  filter(release_year == 2001) %>% 
  arrange(desc(budget))
```

```{r}
LHS %>% RHS
```

```{r}
movies_sel %>% summary()
summary(movies_sel)
```

What movie had the highest vote average...

* overall?
* for movies released after 2010?
* for movies with a vote count greater than the median after 2010?
* for movies that made at least double of their budget?

```{r}
movies_sel %>% 
  arrange(desc(vote_average))
```

```{r}
movies_sel %>% 
  filter(vote_average == max(vote_average))
```

```{r}
movies_sel %>% 
  filter(release_year > 2010,
         vote_average == max(vote_average))
```
...for movies with a vote count greater than the median after 2010?

```{r}
movies_sel %>% 
  filter(vote_count > median(vote_count)) %>% 
  arrange(desc(vote_average))
```
..for movies that made at least double of their budget?

```{r}
movies_sel %>% 
  # Remove anything with a budget of 0 because it skews data
  filter(budget > 0) %>% 
  
  # Calculated percentage change of budget
  mutate(revenue_percent = (revenue - budget) / budget) %>% 
  filter(revenue_percent > 2) %>% 
  arrange(desc(vote_average))
```

How many movies are represented from each language? What language has the second most movies represented?

```{r}
movies_sel %>% 
  group_by(original_language) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  arrange(desc(count))
```
How many movies were released each year?

```{r}
movies_sel %>% 
  group_by(release_year) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  arrange(desc(count))
```

```{r}
movies_sel %>% 
  count(release_year, name = "count", sort = T)
```

How much total budget was used across each original language?

```{r}
movies_sel %>% 
  group_by(original_language) %>% 
  summarize(budget_total = sum(budget)) %>% 
  ungroup() %>% 
  arrange(desc(budget_total))
```

How did vote averages change for English language films over time?

```{r}
movies_sel %>% 
  filter(original_language == "en") %>% 
  group_by(release_year) %>% 
  summarize(vote_average = mean(vote_average, na.rm = T)) %>% 
  ungroup() %>% 
  ggplot(aes(x = release_year,
             y = vote_average)) + 
  geom_line()
```

How many movie titles match their original title? How many did not?

```{r}
movies_sel %>% 
  mutate(title_match_flag = title == original_title) %>% 
  count(title_match_flag) %>% 
  ggplot(aes(x = n,
             y = title_match_flag)) +
  geom_col()
```
How many movie titles match their original title by year?

```{r}
movies_sel %>% 
  mutate(title_match_flag = title == original_title) %>% 
  count(title_match_flag, release_year) %>% 
  ggplot(aes(x = release_year,
             y = n,
             color = title_match_flag)) +
  geom_line()
```

What was the highest budget movie by year?

```{r}
movies_sel %>% 
  group_by(release_year) %>% 
  filter(budget == max(budget)) %>% 
  ungroup() %>% 
  select(release_year, budget, title) %>% 
  arrange(release_year) %>% 
  count(release_year, sort = T)
```

Which title is closest to the median revenue by year?

```{r}
median_revenues_close <- 
  movies_sel %>% 
  group_by(release_year) %>% 
  mutate(revenue_median = median(revenue, na.rm = T),
         revenue_diff = revenue - revenue_median,
         revenue_diff_abs = abs(revenue_diff)) %>% 
  filter(revenue_diff_abs == min(revenue_diff_abs)) %>% 
  ungroup() %>% 
  select(release_year, title, revenue, revenue_median, revenue_diff, revenue_diff_abs) %>% 
  arrange(release_year)
```

```{r}
movies_sel %>% 
  select(id, title, original_language, vote_average) %>% 
  head() %>% 
  dput()
```

```{r}
structure(list(id = c(19995, 285, 206647, 49026, 49529, 559), 
    title = c("Avatar", "Pirates of the Caribbean: At World's End", 
    "Spectre", "The Dark Knight Rises", "John Carter", "Spider-Man 3"
    ), original_language = c("en", "en", "en", "en", "en", "en"
    ), vote_average = c(7.2, 6.9, 6.3, 7.6, 6.1, 5.9)), row.names = c(NA, 
-6L), class = c("tbl_df", "tbl", "data.frame"))
```

```{r}
# install.package("here")
library(here)
```

