---
title: "Week 03 Assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Textbook

-   Chapter 5 - Data transformation
-   Chapter 8 - Workflow: projects
-   Chapter 9 - Introduction
-   Chapter 10 - Tibbles


#### Exercises

We'll be exploring [fuel economy data](https://www.fueleconomy.gov/feg/download.shtml) from 1984-2021. You can review a data dictionary for the dataset [here](https://www.fueleconomy.gov/feg/ws/index.shtml#vehicle).

Utilize the following chunk to get started with your analysis.

```{r}
library(tidyverse)
library(here)

vehicles <- read_csv(here::here("Data", "vehicles.csv"),
                     col_types = cols(.default = "c"))
```

1. Create a reduced version of the dataset that includes the following fields:

-   `make`
-   `model`
-   `year`
-   `cylinders`
-   `drive`
-   `trany`
-   `evMotor`
-   `fuelCost08`
-   `highway08`
-   `city08`
-   `feScore`

Ensure that the `year`, `fuelCost08`, `highway08`, `city08`, and `feScore` fields are coerced to a double.

Assign this to a new dataframe called `vehicles_sub`.

```{r}
vehicles_sub <- 
  vehicles %>% 
  select(make, model, year, cylinders, drive, trany, evMotor, fuelCost08, highway08, city08, feScore) %>% 
  mutate(year = as.double(year),
         fuelCost08 = as.double(fuelCost08),
         highway08 = as.double(highway08),
         city08 = as.double(city08),
         feScore = as.double(feScore))
```

2. Conduct an exploratory analysis of each field in the dataset. This should include:

- Demonstrating the structure of the dataset
- Counts of each character field (in addition to the `year` field)
- Distributions of each numeric field

Indicate anything interesting that you notice.

Hint: When finding counts of fields, sometimes it is helpful to sort the field from highest to lowest values.

```{r}
glimpse(vehicles_sub)
```


```{r}
vehicles_sub %>% 
  count(make, sort = T)

vehicles_sub %>% 
  count(model, sort = T)

vehicles_sub %>% 
  count(year)

vehicles_sub %>% 
  count(cylinders, sort = T)

vehicles_sub %>% 
  count(drive, sort = T)

vehicles_sub %>% 
  count(trany, sort = T)

vehicles_sub %>% 
  count(evMotor, sort = T)
```

```{r}
vehicles_sub %>% 
  ggplot(aes(x = fuelCost08)) +
  geom_histogram()

vehicles_sub %>% 
  ggplot(aes(x = highway08)) +
  geom_histogram()

vehicles_sub %>% 
  ggplot(aes(x = city08)) +
  geom_histogram()

vehicles_sub %>% 
  ggplot(aes(x = feScore)) +
  geom_bar()
```

3. Some of the fields need to be cleaned to ease analysis. 

- Create a new field in `vehicles_sub` called `evMotor_flag` that indicates whether or not the vehicle has an electric motor.

```{r}
vehicles_sub <- 
  vehicles_sub %>% 
  mutate(evMotor_flag = ! is.na(evMotor))
```


4. Create a graph that demonstrates how median city mileage has changed over time.

```{r}
vehicles_sub %>% 
  group_by(year) %>% 
  summarize(city08 = median(city08)) %>% 
  ungroup() %>% 
  ggplot(aes(x = year,
             y = city08)) +
  geom_line()
```




5. Consider the top 5 car makes whose fuel economies have been evaluated most frequently. Create a graph that demonstrates how the city fuel mileage has changed over time for these makes.

```{r}
top_5 <- c("Chevrolet", "Ford", "Dodge", "GMC", "Toyota")

vehicles_sub %>% 
  filter(make %in% top_5) %>% 
  group_by(make, year) %>% 
  summarize(city08 = median(city08)) %>% 
  ungroup() %>% 
  ggplot(aes(x = year,
             y = city08,
             group = make,
             color = make)) +
  geom_line()
```
6. Create a question about the dataset that must be solved by using the `dplyr` verbs - `select`, `mutate`, `summarize`, `arrange`, etc.

-   Indicate what your question is. Consider a question that could be answered best by utilizing facets, multiple geometries, multiple aesthetics, or some combination of those.
-   Create a graph that assists in answering the question.
-   Write a short interpretation of the graph in order to answer your question.



